---
title: "NET Grant Proposal"
author: "Renee Liu"
date: "May 11, 2018"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---


# Lending Club Preliminary Result


```{r message = FALSE, error = FALSE, echo = FALSE}
library(dplyr)
library(mfx)
library(tidyr)
library(plyr)
library(ggplot2)
library(lubridate)
library(stargazer)
```

## What makes a borrower more likely to be rejected by Lending Club? The one with higher loan request?
First, we combine `rejected` and `accepted` loans to see what is the relationship between loan acceptance status and other charateristics associated with a specific loan. We need to run a logistic model to look at the sign on the amount of loan requested.

$$approved status_{i} = \alpha Amount of Loan Request_{i} + \beta fico_{i} + \gamma dti_{i}$$

```{r pooled, echo = FALSE}
#read data
rej = read.csv("rejectedloans.csv")
rej <- rej %>% 
       separate(Application.Date, c("day", "Month", "Year"), sep = "-") %>%
       select(-c(Month, day, Loan.Title, City, State))

#declare logical variable in rej
status = seq(0,0,length.out= nrow(rej))
rej = cbind(rej, status)

appr = read.csv("approvedloans.csv")
appr[, 6] = as.numeric(gsub("%", "", appr[,6]))/100
appr1 <- appr %>%
         na.omit() %>%
         select(issue_d, int_rate, loan_amnt, funded_amnt_inv, term, grade, sub_grade, emp_length, home_ownership, annual_inc, dti, fico_range_low, fico_range_high, inq_last_6mths, open_acc, revol_util) %>%
         mutate(fico = (fico_range_low + fico_range_high)/2) %>%
         separate(issue_d, c("month", "year"), sep = ",") 

for(level in unique(appr1$emp_length)){
  appr1[paste("dummy", level, sep = "_")] <- ifelse(appr1$emp_length == level, 1, 0)
} # Creat dummy variable for employment length

apprj <- appr1 %>%
         select(loan_amnt, year, fico, dti, emp_length)

approved_status = seq(1,1,length.out= nrow(apprj))
apprj = cbind(apprj, approved_status)

names(rej) <- colnames(apprj)
rej$dti = as.numeric(sub("%", "", rej$dti))
pooled = rbind(apprj, rej)
pooled <- pooled %>% drop_na()
pooled$approved_status <- as.factor(pooled$approved_status)
pooled <- droplevels(pooled[!pooled$emp_length == 'n/a',])
pooled_drop1 <- droplevels(pooled[!pooled$emp_length == '< 1 year',])
pooled_dropped_both <- droplevels(pooled_drop1[!pooled_drop1$emp_length == '10+ years',])

#pooled_dropped_both resulted from dropped less than 1 year and more than 10 years emp_length
#not a good way though
pooled_dropped_both$emp_length <- as.character(pooled_dropped_both$emp_length)

p1 <- pooled_dropped_both %>%
  separate(emp_length, c("emp_length", "del") , sep = " ") %>%
  select(-del)

p1$emp_length <- as.numeric(p1$emp_length)
p1$emp_length <- log(p1$emp_length)
p1$loan_amnt <- log(p1$loan_amnt)
p1$fico <- log(p1$fico)

#from now on use p1 to do logistic regression

#pooled$year = year(pooled$year, origin = "2000-01-01")
#pooled$year = as.character.Date(pooled$year)
#pooled$year = year(pooled$year)
#pooled$year = as.numeric(pooled$year)
#pooled$year[pooled$year < 30] = pooled$year + 2000


#fit a logistic regression to test H_0: Higher Amount of Loan request is more likely to be rejected
#Ignoring employment length first
approval_decision_model = glm(approved_status ~ loan_amnt + fico + dti + emp_length, data = p1, family = binomial(link = "logit"))
#stargazer(approval_decision_model, title = "Logistic Regression Results", align = TRUE)
# This produce log odds ratio instead of the direct coef outputs
logit.or = exp(coef(approval_decision_model))
stargazer(approval_decision_model, title = "Logistic Regression Results (log odds ratio)", coef=list(logit.or), p.auto=FALSE, out="logitor.htm")

summary(approval_decision_model)

#summary(approval_decision_model)$coefficients
#alpha = summary(approval_decision_model)$coefficients[2,1]

#Mcfadden's pseudo R-squared
ll.null <- approval_decision_model$null.deviance/-2
ll.proposed <- approval_decision_model$deviance/-2
pseudo_R_2 = (ll.null - ll.proposed) / ll.null

#P-value for pseudo R squared
1 - pchisq(2*(ll.proposed - ll.null),df = (length(approval_decision_model$coefficients)-1))

# marginal effects
logitmfx(approved_status ~ loan_amnt + fico + dti + emp_length, data = p1)

```

```{r}
# Define function - multiplot
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

We see that the coefficient on loan amount reqeusted is `r alpha`, and it is negative with statistical significance.
Within the rejected data, we see that most of 
```{r}
g_rej <- ggplot(rej, aes(emp_length))
emp_rej = g_rej + geom_bar() + scale_y_continuous(labels=function(n){format(n, scientific = FALSE)}) + xlab("Employment Length (Rejected Loans)")
g_apprj <- ggplot(apprj, aes(emp_length))
emp_apprj = g_apprj + geom_bar() + xlab("Employment Length (Approved Loans)")
multiplot(emp_rej, emp_apprj, cols=1)
```
The left and right panel shows the distribution of employment length from Rejected and Accepted data. It shows that most of the people with less-than-one-year of working got screened out by LC. This shows that LC cares a lot about borrowers' qualities.

```{r}
# fico density plots from rejected and approved loans 
# rejected

rejf = ggplot(subset(rej, !is.na(fico)), aes(x = fico))
a = rejf + geom_density(fill = "white") + geom_vline(aes(xintercept = mean(fico)), 
                                     linetype = "dashed", size = 0.6) + xlab("Fico Scores (Rejected Loans)")
# accepted
accf = ggplot(subset(apprj, !is.na(fico)), aes(x = fico))
b = accf + geom_density(fill = "lightgray") + geom_vline(aes(xintercept = mean(fico)), 
                                       linetype = "dashed", size = 0.6) + xlab("Fico Scores (Approved Loans)")
multiplot(a, b, cols=1)


```